How economic, health and social factors affect life expectancy in developing countries?

### 2025-04-10 paper.Rmd
added to sec 5

# Generate a nicely formatted table
kable(model_table, digits = 4, align = "c",
      col.names = c("Variable", "Estimate", "Std. Error", "Pr(>|t|)"))

@@

>>>>

# Generate a nicely formatted table
kable(model_table, digits = 4, align = "c",
      col.names = c("Variable", "Estimate", "Std. Error", "Pr(>|t|)"))
```

@@

Interestingly, our model revealed an unexpected negative coefficient (-1.511) associated with countries categorized as having negligible alcohol consumption, indicating that these countries, on average, have approximately 1.51 years lower life expectancy compared to those with regular alcohol consumption, assuming all other variables are held constant. This counterintuitive result contrasts with the existing literature, where higher alcohol consumption is typically linked to adverse health outcomes and lower longevity (Roffia et al., 2023). It suggests that in developing countries, negligible alcohol consumption may reflect broader socio-cultural, economic, or policy contexts, such as restrictive laws or limited social freedoms, potentially indirectly correlating with restricted healthcare access or lower socioeconomic status. This finding indicates a need for deeper qualitative investigation to understand the broader social determinants involved and their indirect effects on life expectancy.

It is important to note a deviation from previous literature regarding GDP per capita, which was ultimately excluded from our final model due to statistical insignificance. Prior studies consistently demonstrated GDP per capita as a strong predictor positively associated with life expectancy (Miladinov, 2020; Adebayo et al., 2024). However, in our model, the relationship between GDP per capita and life expectancy might have been captured indirectly through other explanatory variables like schooling and healthcare expenditure. These variables may effectively mediate the economic benefits usually associated with GDP, highlighting the importance of explicitly identifying and understanding such mediating mechanisms in developing countries.
